Exploratory: Optimal Resolution

### 2021-04-29 exploratory/exploratory.Rmd
ms updates

for( m in models ){
  phylum <- read_csv(paste0("../data/process/combined-phylum-",m,".csv")) %>% 
    mutate(level="phylum")
  class <- read_csv(paste0("../data/process/combined-class-",m,".csv")) %>% 
    mutate(level="class")
  order <- read_csv(paste0("../data/process/combined-order-",m,".csv")) %>% 
    mutate(level="order")
  family <- read_csv(paste0("../data/process/combined-family-",m,".csv")) %>% 
    mutate(level="family")
  genus <- read_csv(paste0("../data/process/combined-genus-",m,".csv")) %>%
    mutate(level="genus")
  otu <- read_csv(paste0("../data/process/combined-otu-",m,".csv")) %>%
    mutate(level="otu")
  asv <- read_csv(paste0("../data/process/combined-asv-",m,".csv")) %>%
    mutate(level="asv")

@@

  cv_med <- m_auc.df %>%
    filter(auc_type == "cv_metric_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC))

  t_med <- m_auc.df %>% 
    filter(auc_type == "test_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC))

@@

#compare two models levels within a taxonomic level
p.table.level <- read_csv("../analysis/pvalues_by_level.csv")
#compare two taxonomic levels within a model type
p.table.model <- read_csv("../analysis/pvalues_by_model.csv")

@@

## Table of Contents
- [Input data overview](#input-data-overview)
- [Hyperparameter Performance](#hyperparameter-performance)
- [Model Performance](#model-performance)
- [DADA2 Comparison](#dada2-comparison)

@@

input <- read_csv("../analysis/input_values.csv")

@@

ggplot(df_by_method,aes(x=factor(level,
                            levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
                   y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    facet_grid(.~method) + 
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_method.png")
```

@@

ggplot(df_by_level,aes(x=method, y=AUC, fill=auc_type)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(notch = TRUE) + 
  facet_wrap(.~level,nrow = 2) + 
  theme_bw() + xlab("") + ylab("AUROC") + 
  theme(legend.position ="top",
        axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_tax.png")

@@

df_all %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")
ggsave("./figures/all_model_level.png")
```

@@

#mothur and dada2 input values
input_data <- bind_rows(read_csv("../analysis/input_values.csv"),
                        read_csv("../data/dada2/process/summary_input_values.csv"))

@@

dada2_df <- NULL
dada2_auc_df <- NULL
for( m in models ){
  asv <- read_csv(paste0("../data/dada2/process/combined-dada2-",m,".csv")) %>%
    mutate(type="DADA2")
  dada2_df <- bind_rows(dada2_df,asv)

@@

mod_md_df %>% 
  filter(type=="DADA2") %>% 
  ggplot(aes(x=method, y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    ggtitle("DADA2 ASV Model Performance across Models")

mod_md_df %>% 
  ggplot(aes(x=type, y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    facet_grid(.~method) +
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    ggtitle("DADA2 vs Mothur ASV Model Performance")

@@

mod_md_auc_df %>% 
  filter(type=="DADA2") %>% 
  ggplot(aes(x=method,y=AUC)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(aes(fill=method),notch = T) +
    scale_fill_manual(values = pal2) + 
    scale_y_continuous(limits=c(0.2,1.05)) +
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="none",
          axis.text.x = element_text(angle = 45,hjust = 1),
          panel.grid.major.x = element_blank(),
          axis.text = element_text(size=14)) +
    ggtitle("DADA2") +
    stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
    stat_pvalue_manual(sub_stats_level,label="sig")

@@

mod_md_auc_df %>%
  rename(group=type,model=method) %>% 
  ggplot(aes(x=group,y=AUC,color=group)) +
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_jitter() +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    #geom_boxplot(aes(fill=model),notch = T) +
    scale_y_continuous(limits=c(0.4,0.9)) +
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="none",
          axis.text.x = element_text(angle = 45,hjust = 1),
          panel.grid.major.x = element_blank(),
          axis.text = element_text(size=14)) +
    ggtitle("DADA2 v Mothur ASVs") +
    facet_grid(~model) +
    stat_pvalue_manual(sub_stats_model,label="p_value")

@@

```{r dada2_all, echo=F, message=F, out.width="100%",fig.height=3}
dada2_auc_df %>% 
  rename(level=type) %>% 
  bind_rows(auc_df) %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv","DADA2"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV","DADA2")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(notch = TRUE) + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

@@

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/glmnet_with_stats.png",width = 5,height=4,units="in")

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method),notch = T) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rpart2_with_stats.png",width = 5,height=4,units="in")

@@

@@

## Feature Importance
The mikropml packages includes an option for finding [feature importance](http://www.schlosslab.org/mikropml/reference/get_feature_importance.html) using a permutation method. This significantly increases the time to run the models. The function permutes each model feature (e.g. genus) and recalulates the AUC with that feature permuted. The outputs are the permuted AUC value and the difference between this AUC value and the actual AUC value. Larger positive differences can be interpreted as being more important since the AUC value is smaller when this feature is permuted.

@@

>>>>

for( m in models ){
  phylum <- read_csv(paste0("../data/process/combined-phylum-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>% 
    mutate(level="phylum")
  class <- read_csv(paste0("../data/process/combined-class-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>% 
    mutate(level="class")
  order <- read_csv(paste0("../data/process/combined-order-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>% 
    mutate(level="order")
  family <- read_csv(paste0("../data/process/combined-family-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>% 
    mutate(level="family")
  genus <- read_csv(paste0("../data/process/combined-genus-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>%
    mutate(level="genus")
  otu <- read_csv(paste0("../data/process/combined-otu-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>%
    mutate(level="otu")
  asv <- read_csv(paste0("../data/process/combined-asv-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>%
    mutate(level="asv")

@@

  cv_med <- m_auc.df %>%
    filter(auc_type == "cv_metric_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC),.groups = 'drop')

  t_med <- m_auc.df %>% 
    filter(auc_type == "test_AUC") %>%
    group_by(level) %>% 
    summarise(!!m := median(AUC),.groups = 'drop')

@@

#compare two models levels within a taxonomic level
p.table.level <- read_csv("../analysis/pvalues_by_level.csv",
                     col_types = c(p_value = col_double(),.default=col_character()))
#compare two taxonomic levels within a model type
p.table.model <- read_csv("../analysis/pvalues_by_model.csv",
                     col_types = c(p_value = col_double(),.default=col_character()))

@@

## Table of Contents
- [Input data overview](#input-data-overview)
- [Hyperparameter Performance](#hyperparameter-performance)
- [Model Performance](#model-performance)
  - [Random Forest](#random-forest-model-performance-with-significance)
- [DADA2 Comparison](#dada2-comparison)
- [Feature Importance](#feature-importance)

@@

input <- read_csv("../analysis/input_values.csv",
                  col_types = c(level = col_character(), .default = col_double()))

@@

ggplot(df_by_method,aes(x=factor(level,
                            levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
                   y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    facet_grid(.~method) + 
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_method.png")
```

@@

ggplot(df_by_level,aes(x=method, y=AUC, fill=auc_type)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot() + 
  facet_wrap(.~level,nrow = 2) + 
  theme_bw() + xlab("") + ylab("AUROC") + 
  theme(legend.position ="top",
        axis.text.x = element_text(angle = 45,hjust = 1)) 
ggsave("./figures/performance_by_tax.png")

@@

df_all %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(alpha=0.9) + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size=14)) +
    scale_fill_manual(values=pal2,name="Model") #+
  #guides(fill=guide_legend(nrow=2))
ggsave("./figures/all_model_level.png",width = 7,height=4,units="in")
```

@@

#mothur and dada2 input values
input_data <- bind_rows(read_csv("../analysis/input_values.csv",
                                 col_types = c(level = col_character(), .default = col_double())),
                        read_csv("../data/dada2/process/summary_input_values.csv",
                                 col_types = c(level = col_character(), .default = col_double())))

@@

dada2_df <- NULL
dada2_auc_df <- NULL
for( m in models ){
  asv <- read_csv(paste0("../data/dada2/process/combined-dada2-",m,".csv"),
                     col_types = c(method = col_character(),.default=col_double())) %>%
    mutate(type="DADA2")
  dada2_df <- bind_rows(dada2_df,asv)

@@

mod_md_df %>% 
  filter(type=="DADA2") %>% 
  ggplot(aes(x=method, y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    ggtitle("DADA2 ASV Model Performance across Models")

mod_md_df %>% 
  ggplot(aes(x=type, y=AUC, fill=auc_type)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    facet_grid(.~method) +
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    ggtitle("DADA2 vs Mothur ASV Model Performance")

@@

mod_md_auc_df %>% 
  filter(type=="DADA2") %>% 
  ggplot(aes(x=method,y=AUC)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot(aes(fill=method)) +
    scale_fill_manual(values = pal2) + 
    scale_y_continuous(limits=c(0.2,1.05)) +
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="none",
          axis.text.x = element_text(angle = 45,hjust = 1),
          panel.grid.major.x = element_blank(),
          axis.text = element_text(size=14)) +
    ggtitle("DADA2") +
    stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
    stat_pvalue_manual(sub_stats_level,label="sig")

@@

mod_md_auc_df %>%
  rename(group=type,model=method) %>% 
  ggplot(aes(x=group,y=AUC,color=group)) +
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_jitter() +
    stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
    #geom_boxplot(aes(fill=model)) +
    scale_y_continuous(limits=c(0.4,0.9)) +
    theme_bw() + xlab("") + ylab("AUROC") +
    theme(legend.position ="none",
          axis.text.x = element_text(angle = 45,hjust = 1),
          panel.grid.major.x = element_blank(),
          axis.text = element_text(size=14)) +
    ggtitle("DADA2 v Mothur ASVs") +
    facet_grid(~model) +
    stat_pvalue_manual(sub_stats_model,label="p_value")

@@

```{r dada2_all, echo=F, message=F, out.width="100%",fig.height=3}
dada2_auc_df %>% 
  rename(level=type) %>% 
  bind_rows(auc_df) %>% 
  mutate(level=factor(level,levels=c("phylum","class","order","family","genus","otu","asv","DADA2"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV","DADA2")),
         method=factor(method,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(auc_type == "test_AUC") %>% 
  ggplot(aes(x=level,y=AUC,fill=method)) + 
    geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
    geom_boxplot() + 
    theme_bw() + xlab("") + ylab("AUROC") + 
    theme(legend.position ="top",
          axis.text.x = element_text(angle = 45,hjust = 1)) +
    scale_fill_manual(values=pal2,name="Model")

```{r dada2_rf_wstat}

# #subset to just test data
# mod_md_auc_df <- md_auc_df %>% 
#   filter(auc_type=="test_AUC",method=="rf") %>% 
#   mutate(method=factor(method,levels=c("rf"),
#                        labels=c("Random Forest")))
# 
# #compare different models with DADA2 ASVs
# sub_stats_dada2_rf <- dada2_p_model %>% 
#   filter(model=="rf") %>%    
#   mutate(sig=case_when(p_value >= 0.05 ~ "NS",
#                        TRUE ~ "*")) %>% 
#   mutate(y.position=seq.int(0.85,1.05,length.out = nrow(.))) %>% 
#   rename(group1=level1,group2=level2) %>% 
#   filter(sig=="NS")
# 
# mod_md_auc_df %>% 
#   filter(type=="DADA2") %>% 
#   ggplot(aes(x=method,y=AUC)) + 
#     geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
#     geom_boxplot(aes(fill=method)) +
#     scale_fill_manual(values = pal2) + 
#     scale_y_continuous(limits=c(0.2,1.05)) +
#     theme_bw() + xlab("") + ylab("AUROC") +
#     theme(legend.position ="none",
#           axis.text.x = element_text(angle = 45,hjust = 1),
#           panel.grid.major.x = element_blank(),
#           axis.text = element_text(size=14)) +
#     ggtitle("DADA2") +
#     stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
#     stat_pvalue_manual(sub_stats_level,label="sig")
# 
# 
# #compare Mother and DADA2 ASVs across models
# sub_stats_model <- dada2_p_model %>% 
#    mutate(level1=factor(level1,levels=c("asv","dada2"),
#                             labels=c("Mothur","DADA2")),
#           level2=factor(level2,levels=c("asv","dada2"),
#                             labels=c("Mothur","DADA2")),
#          model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
#                        labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
#   mutate(sig=case_when(p_value >= 0.05 ~ "NS",
#                        TRUE ~ "*")) %>% 
#   #filter(sig=="NS") %>% 
#   mutate(y.position=rep(0.85,length.out = nrow(.))) %>% 
#   rename(group1=level1,group2=level2) %>% 
#   mutate(p_value=case_when(p_value < 0.001 ~ round(p_value,digits=4),
#                            TRUE ~ round(p_value,digits=2)))
# 
# mod_md_auc_df %>%
#   rename(group=type,model=method) %>% 
#   ggplot(aes(x=group,y=AUC,color=group)) +
#     geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
#     geom_jitter() +
#     stat_summary(fun.data=median_hilow, fun.args=0.5, geom="crossbar", width=0.5, color="black") +
#     #geom_boxplot(aes(fill=model)) +
#     scale_y_continuous(limits=c(0.4,0.9)) +
#     theme_bw() + xlab("") + ylab("AUROC") +
#     theme(legend.position ="none",
#           axis.text.x = element_text(angle = 45,hjust = 1),
#           panel.grid.major.x = element_blank(),
#           axis.text = element_text(size=14)) +
#     ggtitle("DADA2 v Mothur ASVs") +
#     facet_grid(~model) +
#     stat_pvalue_manual(sub_stats_model,label="p_value")

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

@@

```{r rf_wstat_side,include=F}
sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,.95,length.out = nrow(sub_stats)))
  mutate(y.position=c(0.85,0.95,0.9,.95))

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")),
                           y=AUC)) + 
  #geom_jitter(aes(color=method,width=0,height=0.2)) +
  geom_boxplot(aes(fill=method,alpha=0.8)) +
  scale_fill_manual(values = pal2[m]) +
  #scale_color_manual(values = pal2[m]) +
  #scale_x_continuous(limits=c(0.4,1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_pvalue_manual(sub_stats,label="sig",coord.flip = T,angle = 0,hjust=0) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") 
ggsave("./figures/rf_with_stats_sideways.png",width = 5,height=4,units="in")
```

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/glmnet_with_stats.png",width = 5,height=4,units="in")

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rpart2_with_stats.png",width = 5,height=4,units="in")

@@

<div class = "row">
<div class = "col-md-6">
```{r svmRadial_table,echo=FALSE}
t_table %>%
  select(level,svmRadial) %>% 
  mutate(svmRadial=round(svmRadial,digits = 3)) %>% 
  rename("Median AUC"=svmRadial,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r svmRadial_plot,echo=F,message=F}
## svmRadial Model Performance
m <- "SVM Radial"

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/svmRadial_with_stats.png",width = 5,height=4,units="in")

@@

<div class = "row">
<div class = "col-md-6">
```{r xgbTree_table,echo=FALSE}
t_table %>%
  select(level,xgbTree) %>% 
  mutate(xgbTree=round(xgbTree,digits = 3)) %>% 
  rename("Median AUC"=xgbTree,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r xgbTree_plot,echo=F,message=F}
## XGBoost Model Performance
m <- "XGBoost"

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/xgbTree_with_stats.png",width = 5,height=4,units="in")

## Feature Importance
The mikropml packages includes an option for finding [feature importance](http://www.schlosslab.org/mikropml/reference/get_feature_importance.html) using a permutation method. This significantly increases the time to run the models. The function permutes each model feature (e.g. genus) and recalculates the AUC with that feature permuted. The outputs are the permuted AUC value and the difference between this AUC value and the actual AUC value. Larger positive differences can be interpreted as being more important since the AUC value is smaller when this feature is permuted.

@@

```{r imp_plots}
imp_plots <- list()
m="rf"
for(l in levels){
  test <- read_csv(paste0("../data/process/importance-",l,"-",m,".csv"),
                   col_types = c(perf_metric = col_double(),
                                 perf_metric_diff = col_double(),
                                 seed = col_double(),
                                 .default = col_character())) %>% 
    mutate(level := !!l)
  test_results <- read_csv(paste0("../data/process/combined-",l,"-",m,".csv"),
                           col_types = c(method = col_character(), .default = col_double())) %>% 
    mutate(level := !!l)

  test_tax <- read_tsv(paste0("../data/",l,"/crc.taxonomy"),
                   col_types = c(Size = col_double(),
                                 .default = col_character())) %>%
    mutate(level := !!l)

  test_merge <- test %>%
    select(perf_metric,perf_metric_diff,names,method,seed,level) %>% 
    rename(AUC=perf_metric,OTU=names,AUC_diff=perf_metric_diff) %>% 
    inner_join(test_tax,by=c("level","OTU")) %>% 
    mutate(name=paste0(Taxonomy," (",OTU,")"))

  test_quartiles <- quantile(pull(test_results,AUC))
  names(test_quartiles) <- c("q0","q1","q2","q3","q4")

  test_median <- median(pull(test_results,AUC))

  if(l %in% c("otu","asv")){
    imp_plots[[paste0(m,"-",l,"-1")]] <- test_merge %>% 
      mutate(OTU=reorder(OTU,desc(AUC),.fun="median")) %>% 
      ggplot(aes(x=AUC,y=OTU)) + 
        geom_point(alpha=0.5,color="lightblue") +
        stat_summary(fun = mean,geom = "point", shape = 18, size = 3,aes(color="Mean")) +
        stat_summary(fun = median,geom = "point", shape = 18, size = 3,aes(color="Median")) +
        scale_color_manual(name="",values=c("red","green")) +
        theme_bw() +
        theme(legend.position = "bottom",
              axis.text.y = element_blank()) +
        geom_vline(xintercept = test_quartiles["q1"]) +
        geom_vline(xintercept = test_quartiles["q3"]) + 
        geom_vline(xintercept = test_median,lty="dashed") +
        ggtitle(paste0(m,"-",l,": AUC ordered")) +
        xlab("Permuted AUC") + ylab("Feature")
    #ggsave(paste0("./data/analysis/exploratory3/importance_metric_",m,"_",l,".png"))

  imp_plots[[paste0(m,"-",l,"-2")]] <- test_merge %>% 
    mutate(OTU=reorder(OTU,AUC_diff,.fun="median")) %>% 
    ggplot(aes(x=AUC_diff,y=OTU)) + 
      geom_point(alpha=0.5,color="lightblue") +
      stat_summary(fun = mean,geom = "point", shape = 18, size = 3,aes(color="Mean")) +
      stat_summary(fun = median,geom = "point", shape = 18, size = 3,aes(color="Median")) +
      scale_color_manual(name="",values=c("red","green")) +
      geom_vline(xintercept = 0,lty="dashed",color="black") +
      theme_bw() +
      theme(legend.position = "bottom",
            axis.text.y = element_blank()) +
      ggtitle(paste0(m,"-",l,": AUC difference sorted")) +
      ylab("Feature") + xlab("AUC Difference (actual - permuted)")
    #ggsave(paste0("./data/analysis/exploratory3/importance_diff_",m,"_",l,".png"))

  }else{

    imp_plots[[paste0(m,"-",l,"-1")]] <- test_merge %>% 
      mutate(name=reorder(name,desc(AUC),.fun="median")) %>% 
      ggplot(aes(x=AUC,y=name)) + 
        geom_point(alpha=0.5,color="lightblue") +
        stat_summary(fun = mean,geom = "point", shape = 18, size = 3,aes(color="Mean")) +
        stat_summary(fun = median,geom = "point", shape = 18, size = 3,aes(color="Median")) +
        scale_color_manual(name="",values=c("red","green")) +
        theme_bw() +
        theme(legend.position = "bottom") +
        geom_vline(xintercept = test_quartiles["q1"]) +
        geom_vline(xintercept = test_quartiles["q3"]) + 
        geom_vline(xintercept = test_median,lty="dashed") +
        ggtitle(paste0(m,"-",l,": AUC ordered")) +
        xlab("Permuted AUC") + ylab("Feature")
    #ggsave(paste0("./data/analysis/exploratory3/importance_metric_",m,"_",l,".png"))

    imp_plots[[paste0(m,"-",l,"-2")]] <- test_merge %>% 
      mutate(name=reorder(name,AUC_diff,.fun="median")) %>% 
      ggplot(aes(x=AUC_diff,y=name)) + 
        geom_point(alpha=0.5,color="lightblue") +
        stat_summary(fun = mean,geom = "point", shape = 18, size = 3,aes(color="Mean")) +
        stat_summary(fun = median,geom = "point", shape = 18, size = 3,aes(color="Median")) +
        scale_color_manual(name="",values=c("red","green")) +
        geom_vline(xintercept = 0,lty="dashed",color="black") +
        theme_bw() +
        theme(legend.position = "bottom") +
        ggtitle(paste0(m,"-",l,": AUC difference sorted")) +
        ylab("Feature") + xlab("AUC Difference (actual - permuted)")
    #ggsave(paste0("./data/analysis/exploratory3/importance_diff_",m,"_",l,".png"))

  }
}

```{r rf_importance_plots,echo=F,message=F,fig.show="hold",out.width="50%"}
imp_plots[["rf-phylum-1"]]
imp_plots[["rf-phylum-2"]]

imp_plots[["rf-class-1"]]
imp_plots[["rf-class-2"]]

imp_plots[["rf-order-1"]]
imp_plots[["rf-order-2"]]

imp_plots[["rf-family-1"]]
imp_plots[["rf-family-2"]]

imp_plots[["rf-genus-1"]]
imp_plots[["rf-genus-2"]]

### Top 10 Important features
```{r importance_rank, message=F,echo=F}
